---
title: "Chapter 1"
author: "chenshuai"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

## Q1

(a) 

$Pr(y) = Pr(\theta = 1) \cdot N(1,2^2) + Pr(\theta = 2) \cdot N(2, 2^2)$

```{r}
y = seq(-5, 10, 0.01)
p1 = dnorm(y, 1, 2)
p2 = dnorm(y, 2, 2)
p = 0.5 * p1 + 0.5 * p2
plot(y, p, type = "l")
```

(b) 

$Pr(\theta=1|y=1) = \frac{Pr(\theta=1,y=1)}{Pr(y=1)}=\frac{Pr(\theta = 1) \cdot N(1|1,2^2)}{Pr(\theta = 1) \cdot N(1|1,2^2) + Pr(\theta = 2) \cdot N(1|2, 2^2)} = 0.53$

```{r}
numerator = 0.5 * dnorm(1, 1, 2)
denomnator = 0.5 * dnorm(1, 1, 2) + 0.5 * dnorm(1, 2, 2)
Pr = numerator / denomnator
```

(c) 

```{r}

posterior <- function(sigma){
  numerator1 = 0.5 * dnorm(1, 1, sigma)
  numeroator2 = 0.5 * dnorm(1, 2, sigma)
  denomnator = 0.5 * dnorm(1, 1, sigma) + 0.5 * dnorm(1, 2, sigma)
  Pr = c(numerator1 / denomnator, numeroator2 / denomnator)
  return (Pr)
}

data <- data.frame(
  Category = c("theta = 1", "theta = 2"),
  Magnitude = posterior(0.5)
)
p1 <- data |>  ggplot(aes(x = Category, y = Magnitude)) + geom_bar(stat = "identity", fill = "steelblue",width = .5) + labs(title = "sigma = 0.5", x = "theta",y="density") + ylim(0,1)+ theme_minimal()

data <- data.frame(
  Category = c("theta = 1", "theta = 2"),
  Magnitude = posterior(2)
)
p2 <- data |>  ggplot(aes(x = Category, y = Magnitude)) + geom_bar(stat = "identity", fill = "red",width = .5) + labs(title = "sigma = 2", x = "theta",y="density") + ylim(0,1)+theme_minimal()


grid.arrange(p1, p2, nrow = 1)

```

## Q2

For the ith element in the vector $u$, $u_i$ is a random variable, and $p(u_i|v)$ is the probability density function of $u_i$ given $v$. We also have $E(u_i)=\int \int u_ip(u_i,v)du_idv=\int \int u_ip(u_i|v)p(v)du_idv=\int\int u_ip(u_i|v)du_ip(v)dv=\int E(u_i|v)p(v)dv = E(E(u_i|v))$

For diagonal elements of $var(u_i)$, (1.9) states that $var(u_i) = E(var(u_i|v))+ var(E(u_i|v))$. For off-diagonal elements, we want to prove $cov(u_i,u_j)=E[cov(u_i, u_j|v)] + cov[E(u_i|v), E(u_j|v)]$.

First we expand the RHS,

The first term :$cov(u_i, u_j|v) = E(u_iu_j|v) - E(u_i|v)E(u_j|v)$

The second term:$cov(E(u_i|v), E(u_j|v))=E(E(u_i|v)E(u_j)|v)-E(E(u_i|v))E(E(u_j|v))$

Then we have $E(E(u_i|v)E(u_j)|v)$ in both terms cancelled out, and leave

$E(E(u_iu_j|v)) - E(E(u_i|v))E(E(u_j|v)) = E(u_iu_j)-E(u_i)E(u_j)=cov(u_i,u_j)$

## Q3

We have$Pr(xx) = p,Pr(xX)=2p(1-p),Pr(XX)=(1-p)^2$,

The probability of brown eyes in population is $Pr(brown) = Pr(xx) + Pr(xX)$

The joint probability that brown eye parents have brown eye children is $0.75\cdot Pr(xX) \cdot Pr(xX)+Pr(xX) \cdot Pr(XX)+Pr(XX)\cdot Pr(XX)\\=0.75 \cdot 2p(1-p) \cdot 2p(1-p) +2p(1-p)(1-p)^2+(1-p)^2(1-p)^2$, and among those children, the probability that they are homozygous is $0.5\cdot Pr(xX) \cdot Pr(xX)+0.5\cdot Pr(xX) \cdot Pr(XX)$

So the prior for brown eye kid with brown eye parent is $\frac{2p}{1+2p}$.

Therefore, the probability that Judy is homozygous given that all her children have brown eyes is
Pr(Judy is Xx | n children all have brown eyes) = $\frac{prior \cdot (\frac{3}{4})^n}{prior \cdot (\frac{3}{4})^n+(1-prior)\cdot 1}$, the other term in denomnator means the probability that Judy is XX and all her children are brown eyes, they sum up and give the probability that all her children are brown eyes.



```{r}

## Q4

## Q6

## Q7

## Q8
